Reinforcement \change{L}earning for \change{O}ptions\\on \change{T}arget \change{V}olatility \change{F}unds\thanks{Supported by ISCRA from Cineca, \email{iscra@cineca.it}.}

### 2023-03-21 tex/m2ef/dnpp_referee_response.tex

\noindent {\bf ``Reinforcement learning for options on target volatility funds''.\smallskip\\ Response note to referee's report.}\\

\noindent
We thank the Referee for her interesting comments and all the appreciated suggestions to improve our work. We add few remarks to their comments in the following sections and attach a revised version of the paper. We believe that these extensions address all the points highlighted in the reviews and we hope that they will be considered sufficient to recommend our work for the publication.\\

{\bf $\bullet$ Answers to referee's report}\\

We list here a detailed list of responses to Referee's major remarks. We address the minor remarks directly in the text.

@@

\begin{enumerate}

 \item \textit{[Both in Abtract and Introduction, the presentation of the problem to deal with is not well structured. As a matter of fact, first the TVS is briefly introduces, then options and TVO are only hinted, and lastly the problem is shortly described, without all these topics are well linked each other. I suggest the authors to improve this presentation.]}

 In the revised manuscript, we conduct a comprehensive review of the crucial components of the Abstract and Introduction sections to ensure that the reader is presented with a lucid and concise overview of the problem being addressed. In the Introduction, our focus is primarily on establishing a stronger connection between TVSs and the derivatives market, which are based on these underlying assets. Furthermore, we improve the description of the problem addressed in this paper, with a particular emphasis on the methods employed and the results. We also welcome the referee's suggestion in the minor remarks to improve the title of the manuscript and the related Keywords. We hope that this will enable the reader to contextualize our work more effectively within the pertinent literature.

@@

 \item \textit{[The Authors address the investigated continuous stochastic control problem through two different approaches: an exact one providing the analytical solution of the problem, and an approximate one based on RL. As for the previous point, even in this one the connections between the two solution approaches are not entirely intelligible. In other terms, it is not entirely clear how the two approaches relate each other. Is the approximate one necessary given that the exact approach provides an analytical solution to the problem? What is the "utility" to use an approximate approach given the applicability of an exact one? I suggest the Authors to clarify this issue.]}

The analytical solution applies only for a specific choice of the model for the dynamics of the underlying  assets and under suitable hypotheses on the payoff. We add equation (52) and a sentence in the first paragraph of Section 4 to stress this point. Furthermore, we enhance the clarity of this point in the Introduction section where we discuss the contribution of this work.

@@

 \item \textit{[In Subsection 3.3, three (so-called intuitive) strategies are introduced, $S_A$, $S_B$ and $S_C$. No rationales are provided concerning their choices. I suggest the Author to give some details about this point.]}

 The three \quotations{intuitive strategies} presented in this paper represent plausible expert-based approaches that a practitioner may choose to adopt when addressing the control problem. We provide further elucidation on this matter in Subsection 3.3, where we also incorporate the financial implications of these strategies.

@@

 \item \textit{[In Section 4, the Authors state that ``[t]he first [RL] method is a specific algorithm developed by us to fit the problem [...]''. As know, an important feature of any RL method is its convergence to the (unknown) best policy, if any. The Authors do not spend words about this characteristic of the specific algorithm they developed and applied, so I ask them to give some detail about this topic.]}

 We rephrase the given sentence to clarify that we adopt an existing algorithm well-known in the Reinforcement Learning literature to tackle the problem at hand. In formal terms, it can be classified as a direct policy search algorithm that relies on gradient-based optimization of the Monte Carlo outcome obtained from numerical simulations.

@@

 \item \textit{[In Subsection 5.1, the Author claim that an appropriate choice of the learning rate starting value constitutes ``a good choice in terms of [...] avoiding over-fitting''. As known, when training a FFNN, standard methods for avoiding over-fitting are based on error metrics calculated over validation sets or through $K$-fold validation approaches. So, it is not clear as the Authors can guarantee no over-fitting without exploiting such methods or similar. I suggest the Authors to carefully clarify this issue.]}

We add to the second paragraph of Section 5.1 a sentence which explains that over-fitting is detected on a generated validation set.

### 2023-03-21 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
minor fixes

\institute{Financial Engineering, Intesa Sanpaolo Group,\\
Largo Mattioli 3, 20121 Milan, Italy\\
\email{\{Roberto.Daluiso, Andrea.Pallavicini\}@intesasanpaolo.com} \and
\changeE{Digital Development Sales \& Trading Milan, Unicredit Bank AG Milano,}\\
\changeE{Piazza Gae Aulenti, 4, 20154 Milan, Italy}\\
\email{\changeE{Emanuele.Nastasi@unicredit.eu}} \and \change{Financial Risk Methodologies, Unicredit S.p.A.,\\ Piazza Gae Aulenti, 3, 20154 Milan, Italy \\ \email{Stefano.Polo@unicredit.eu}}
}

@@

We highlight that, to derive the dynamics expressed in Equation~\eqref{eq:TVS_last}, we have not made any assumptions on the risky allocation strategy; thus all this argument is valid for any constraints on the process $\alpha_t$\change{ as, for instance, the case of contracts in which the fund manager is restricted to holding only long positions meaning that $\alpha_t^i>0$ $\forall i = 1,\dots,n$, or situations where the combined total of long and short trades must be capped at a specific contractual amount}.

@@

\subsection{Stochastic \change{o}ptimal \change{c}ontrol \change{p}roblem}
In presence of valuation adjustments, we need to solve the full optimization problem. We discretize the optimal strategy $\alpha_t$ as\footnote{We use the symbol $\mathbf{1}_A$ for the indicator function of a subset $A$.}
\begin{equation}
    \alpha_t \coloneqq \sum_k \mathbf{1}_{ \{t \in [T_{k-1}, T_k)\}}\alpha_{T_{k-1}} \; ,
\label{eq:piecewise_strategy}\end{equation}
according to a time grid $\mathcal{T}\coloneqq \{T_0,...,T_k,...,T_m\}$ with $T_0\coloneqq t$ the pricing date and $T_m \coloneqq T$ the maturity of the option. 
Therefore we can apply the dynamic programming principle to express the optimal $\alpha_t$ at time $T_{k-1}$ as
\begin{equation}
    \alpha_{T_{k-1}}\coloneqq \argmax_{\alpha} \left\{\mathbb{E}_{T_{k-1}}\left[D\left(T_{k-1},T_{k} \right) V_{T_{k}}\left(X_{T_{k}}, I_{T_{k}}(\alpha)\right) \mid X_{T_{k-1}}, I_{T_{k-1}}\right]\right\} \; ,
\label{eq:recursion}\end{equation}
where $V_{T_k}$ is the option value at time $T_k$ and $X$ is any Markovian state such that the drift and the diffusion coefficient of $I_t$ are a function of $\left(X_t,I_t,\alpha_t\right)$ \changeE{and we indicated with $I_{T_k}\left(\alpha\right)$ the value of the strategy on the point $T_{k}$ for a given choice of the weights $\alpha$ at the previous step on the time grid.} We calculate \changeE{the strategy value for the optimal weights $\alpha_{T_{k-1}}$, namely} $I_{T_k}\left(\alpha_{T_{k-1}}\right)$ \removeE{for any given strategy $\alpha_{T_{k-1}}$} by a suitable discretization of \eqref{eq:TVS_last} starting from $X_{T_{k-1}}$ and $I_{T_{k-1}}$\removeE{.} \changeE{and}
\removeE{Thus} \changeE{thus} the derivative price is given by:
\begin{equation}
\begin{aligned}
	    V_{T_{k-1}}\left(X_{T_{k-1}}, I_{T_{k-1}}\right)= \\ 
	     \mathbb{E}_{T_{k-1}}\left[D\left(T_{k-1},T_{k}\right) V_{T_{k}}\left(X_{T_{k}}, I_{T_{k}}\left(\alpha_{T_{k-1}}\right)\right) \mid X_{T_{k-1}}, I_{T_{k-1}}\right] \; ,
\end{aligned}
\label{eq:recursion2}\end{equation}
while the iteration starts from maturity date where the boundary condition is set equal to the payoff function:
\begin{equation}
    V_{T_m} = \Phi\left(I_{T_m}\right) \; .
\end{equation}

@@

\begin{proposition}\label{prop:bs}
When the underlying securities follow a Black and Scholes model with deterministic rates, the optimal strategy for a non-decreasing European payoff consists in minimizing the local drift function, independently of the current state $I_t$
\begin{equation}
    \alpha^*(t) \coloneqq \argmin_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t) \|} \; .
\label{eq:BS_optimal_strategy}
\end{equation}
Analogously, the optimal strategy for a non-increasing European payoff consists in maximizing the local drift function:
\begin{equation}
    \alpha^*(t) \coloneqq \argmax_\alpha \frac{\alpha \cdot \mu(t)}{\norm{\alpha \cdot \nu(t)}} \; .
\label{eq:BS_optimal_strategy_put}
\end{equation}
\end{proposition}
The absence of stochastic elements in Equation~\eqref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori} with no \change{numerical} simulation needed \change{to solve the control problem}; in fact\change{,} \remove{one can solve the optimization problem once for all $\forall t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities} \change{the functions $\mu(t)$ and $\nu(t)$ can be directly derived from the market quotes of the risky assets and of its derivative contracts, then we can apply Equation~\eqref{eq:BS_optimal_strategy} $\forall t \in \mathcal{T}$  to retrieve the deterministic optimal strategy}. Once $\alpha^*$ is known, then one can price the payoff by the following BS formula
\begin{equation}
	V_0^{\text{BS}} = BS(F^{\text{TVS}}(0,T;\alpha^*), K, T, \bar{\sigma}, D(0,T;\zeta)) \; ,
\label{eq:bs_closed_solution}\end{equation}
where $F^{\text{TVS}}(t,T;\alpha)$ is the TVS forward curve defined by
\begin{equation}
	F^{\text{TVS}}(t,T;\alpha) = I_t \exp\left[\int_t^T\left(r(u)-\ell_{\alpha} \left(u\right)\right)du\right] \; ,
\label{eq:tvs_forward}\end{equation}
while $BS(F,K,T,\sigma,D)$ is the standard BS formula for a European option with forward curve $F$, strike $K$, time to maturity $T$, volatility $\sigma$ and discount factor $D$.

In Figure~\ref{fig:strategy_price_comparison} we provide a comparison among the option price obtained with the optimal strategy (BS$^*$) of Equation~\eqref{eq:BS_optimal_strategy} and those with \change{the below expert-based single-asset} \remove{other intuitive} strategies (S$_A$, S$_B$, S$_C$). Here we consider the case of an at-the-money call option with spot $I_0 = 1\, \text{EUR}$, maturity $T=5 \, \text{yr}$, target volatility $\bar{\sigma}=5\%$, and a nonnegative constraint on $\alpha$. The intuitive strategies are
\begin{itemize}
\item S$_A$: invest all in the asset with the maximum forward curve at maturity; \change{this strategy consists of selecting a static portfolio that is never rebalanced, such that the TVO moneyness is maximized;}
\item S$_B$: \remove{for each market pillar before maturity} \change{$\forall t \in \mathcal{T}$} invest all in the asset with minimum $\mu(t)$; \change{in this case we consider a fund manager who rebalances the weights $\alpha(t)$ over the life of the portfolio by selecting the asset with the highest expected return at a given point in time;} 
\item S$_C$: \remove{for each market pillar before maturity} \change{$\forall t \in \mathcal{T}$} invest all in the asset with minimum $\mu(t)/\|\nu(t)\|$\remove{.}\change{; the strategy is akin to S$_B$, with the exception being that the manager endeavors to maximize the expected return that has been adjusted to account for the target volatility mechanism..}
\end{itemize}
As the reader can observe, our strategy outperforms any other \remove{intuitive} \change{expert-based} one a practitioner might adopt \change{to face the control problem}.
\begin{figure}[t]
	\centering
	\includegraphics[width=4in]{strategy_price_comparison_grey.eps}
	\caption{Comparison of plain vanilla prices on the TVS by adopting different allocation strategies: the optimal Black and Scholes solution (BS$^*$) of Equation~\eqref{eq:BS_optimal_strategy} and \change{three expert-based single-asset} strategies (S$_A$, S$_B$, S$_C$)}
	\label{fig:strategy_price_comparison}
\end{figure}

@@

\subsubsection{Active \change{a}sset (or \change{b}ang\change{-b}ang) \change{s}olution}
A closed form solution to the minimization of the local drift correction \eqref{eq:BS_optimal_strategy} can also be derived in the common case that all costs of carry \change{$\mu$} are nonnegative\change{, as they are costs,} and the only constraint on portfolio weights is nonnegativity, which would mean a long-only strategy by the fund manager.

@@

Conversely, if we deal with a dynamic model for which the derivative contract $V$ depends on $X$ then the volatility versor term, namely the second one, in the $\max$ operator of Equation~\eqref{eq:HJB_general} is no longer zero and thus one must solve the entire control problem numerically. \changeB{Indeed, one can check by direct substitution that the natural generalization of \eqref{eq:HJB_solution}
\begin{equation}
 \alpha^*(X_t) = \argmin_\alpha \frac{\alpha \cdot \mu(X_t)}{\|\alpha \cdot \nu(X_t)\|}
\end{equation}
fails to satisfy \eqref{eq:HJB_general}.}

In \removeB{this work}\changeB{what follows} we tackle the non-trivial case of a local volatility \remove{(LV)} model for the $S_t$-dynamics, such that $\nu_t = \nu(t,S_t)$. We have chosen the LV model since it is well known in financial literature and among practitioners \change{\cite{Gatheral2006}}.

\section{Reinforcement learning}\label{sec:RL}
As we have discussed in the previous sections, \changeB{the analytical solution described in Proposition \ref{prop:bs} crucially relies on the assumption that the underlyings evolve according to a Black and Scholes model, and on a monotonicity assumption on the payoff. Therefore,}
one must resort to numerical approaches to solve the stochastic control problem related to the TVS in the case of general payoffs or risky securities dynamics. The standard approach could be to use classical techniques based on \remove{backwards} \change{backward} recursion \eqref{eq:recursion}-\eqref{eq:recursion2} such as American Monte Carlo \cite{Longstaff2001}. However, their performances degrade exponentially as the dimension $n$ of the problem increases, making prohibitively costly finding the solution to the problem. In our contribution, we adopt a novel technique which is free from the curse of dimensionality and is gaining popularity in many scientific branches for solving stochastic optimal control problems: Reinforcement Learning (RL) \cite{Sutton2018}.

\remove{Reinforcement Learning} \change{RL} is a branch of Machine Learning which allows an artificial agent to interact with an environment through actions and observations in order to maximize some notion of cumulative reward \change{called value function}. In RL\change{,} the agent is not told which actions to take but instead, it \remove{must} \change{has to} discover by trial and error which are the behaviors  yielding the highest reward. This is obtained by updating the agent policy $\pi$ which is a mapping from the environment states to the set of actions. Thus RL is independent of pre-collected data as opposed to other Machine Learning techniques\change{, that have to be fed a pre-existing dataset to learn from}. Because of its nature, RL has been successful in quantitative finance for solving control problems; among the most important RL applications in this field, we refer to \cite{Deng2017} as the pioneers in studying self-taught reinforcement trading problems, while to \cite{Halperin2020} and \cite{Kolm2019} for hedging derivatives with RL under market frictions.

@@

In this algorithm, we parameterize the agent policy with a feed forward neural network (FFNN), such that $\theta_j$ coincides with the hidden weights, $s_k$ with the \remove{inputs} \change{input neurons}, and $a_k$ with the output \change{ones}. In this way, we are dealing with a deterministic policy, \change{that maps each state directly to a single action without any addition randomness or probability distribution;} \remove{where its} \change{the map} functional form is given by the neural network. The parameters update is performed as follows: the agent collects a finite batch of experiences interacting with the environment in a set of episodes $\tau$, then the loss-\change{value} function is evaluated as
\begin{equation}
\change{L(\theta_j) =\hat{\mathbb{E}}\left[ \sum_{k=0}^{m-1} r_{k+1} | \theta_j \right]} \; ,
\label{eq:direct_policy_loss}\end{equation}
\begin{equation}
\hcancel[red]{\change{L(\theta_j) = D(0,T;\zeta)\hat{\mathbb{E}}\left[ \Phi(I_T(\alpha(\theta_j))) \right]}} \; ,
\end{equation}  
where the expectation $\hat{\mathbb{E}}[\dots]$ indicates the empirical average on the batch. Then the parameters are updated by plugging the policy into a stochastic gradient ascent algorithm \change{\cite{Sutton2018} with the aim of maximizing Equation~\eqref{eq:direct_policy_loss}}. \remove{The choice of Equation~\eqref{eq:direct_policy_loss} is justified by the fact that, when the algorithm will find the optimal set $\theta^*$, then $L(\theta^*)$ will be a good proxy of the optimal option price.}

@@

\change{As the PPO algorithm is more sophisticated than the previous one, we have decided to test its performance by means of two different reward functions. The former is the same as defined by Equation~\eqref{eq:reward1}, while the latter is given by}
\remove{The second reward function we have defined is}
\begin{equation}
 	r_{k+1} \coloneqq  \gamma^{k}[V_{\text{BS}}(T_{k+1})-V_\text{\text{BS}}(T_k)] \; ,
 \label{eq:reward2}\end{equation}
where $\gamma \in[0,1]$ is an hyper-parameter of the PPO, while $V_{\text{BS}}(T_k)$ is a proxy of the residual option price defined by
\begin{equation}
\begin{aligned}
	V_{\text{BS}}(T_k) &\coloneqq BS(F^{\text{TVS}}(T_k,T;\alpha^*_{\text{BS}},K),K,T-T_k,\bar{\sigma},D(T_k,T;\zeta)) \; ,\\
	V_{\text{BS}}(T_0)&=0 \; , 
\end{aligned}
\end{equation} 
with $\alpha^*_{BS}$ the BS optimal strategy \eqref{eq:BS_optimal_strategy} evaluated at the state $s_k$. In this form the agent actions are hidden inside the term $I_t$ used to compute the TVS forward curve $F^{\text{TVS}}$ defined by Equation~\eqref{eq:tvs_forward}. In this case, the reward function does not suffer of nil values for $0<T_k<T$: the RL agent always gets a feedback from the environment for its choices.  \remove{The} \change{Note that the} hyper-parameter $\gamma$ plays the role of a discount factor in the sense that, as $\gamma$ approaches to zero, the RL agent will tend to maximize immediate rewards while neglecting possible larger rewards in the future.
If we take the cumulative reward per episode and set the PPO parameters\footnote{We refer to \cite{Schulman2016} for a more detailed description for the generalized advantage estimation parameter $\lambda$.} $\gamma =\lambda =1$ we obtain
\begin{equation}
	R(\tau) = \sum_{k=0}^{m-1} r_{k+1} \underset{\gamma=\lambda =1}{=} \sum_{k=0}^{m-1}[V_{\text{BS}}(T_{k+1})-V_\text{\text{BS}}(T_k)] = V_{BS}(T) = (I_T-K)^+ \; ,
\end{equation} 
which is equal to the intrinsic value of the option. This result does not depend on the definition of $V_{\text{BS}}$  $\forall T_k<T$, but we conjecture that the closer $V_{\text{BS}}$ is to the value function, the easier the agent is in learning.

>>>>

\institute{Financial Engineering, Intesa Sanpaolo Group,\\
Largo Mattioli 3, 20121 Milan, Italy\\
\email{\{Roberto.Daluiso, Andrea.Pallavicini\}@intesasanpaolo.com} \and
\change{Capital Market Technologies, swissQuant Group AG,}\\
\change{Kuttelgasse 7, 8001 Z\"urich, Switzerland}\\
\email{\change{Nastasi@swissquant.com}} \and Financial Engineering,
X-Numeris S.r.l.,\\ Via Turro, 7 , 20127 Milan, Italy\\
\change{\email{Sp.Stefano.Polo@gmail.com}}}

@@

We highlight that, to derive the dynamics expressed in Equation~\eqref{eq:TVS_last}, we have not made any assumptions on the risky allocation strategy; thus all this argument is valid for any constraints on the process $\alpha_t$\change{ as, for instance, the case of contracts in which the fund manager is restricted to holding only long positions meaning that $\alpha_t^i\geq0$ $\forall i = 1,\dots,n$, or situations where the combined total of long and short trades must be capped at a specific contractual amount}.

@@

\subsection{Stochastic \change{o}ptimal \change{c}ontrol \change{p}roblem}
In presence of valuation adjustments, we need to solve the full optimization problem. We discretize the optimal strategy $\alpha_t$ as\footnote{We use the symbol $\mathbf{1}_A$ for the indicator function of a subset $A$.}
\begin{equation}
    \alpha_t \coloneqq \sum_k \mathbf{1}_{ \{t \in [T_{k-1}, T_k)\}}\alpha_{T_{k-1}} \; ,
\label{eq:piecewise_strategy}\end{equation}
according to a time grid $\mathcal{T}\coloneqq \{T_0,...,T_k,...,T_m\}$ with $T_0\coloneqq t$ the pricing date and $T_m \coloneqq T$ the maturity of the option. 
Therefore we can apply the dynamic programming principle to express the optimal $\alpha_t$ at time $T_{k-1}$ as
\begin{equation}
    \alpha_{T_{k-1}}\coloneqq \argmax_{\alpha} \left\{\mathbb{E}_{T_{k-1}}\left[D\left(T_{k-1},T_{k} \right) V_{T_{k}}\left(X_{T_{k}}, I_{T_{k}}(\alpha)\right) \mid X_{T_{k-1}}, I_{T_{k-1}}\right]\right\} \; ,
\label{eq:recursion}\end{equation}
where $V_{T_k}$ is the option value at time $T_k$ and $X$ is any Markovian state such that the drift and the diffusion coefficient of $I_t$ are a function of $\left(X_t,I_t,\alpha_t\right)$ \change{and we indicated with $I_{T_k}\left(\alpha\right)$ the value of the strategy at time $T_{k}$ for a given choice of the weights $\alpha$ at time $T_{k-1}$.} We calculate \change{the strategy value for the optimal weights $\alpha_{T_{k-1}}$, namely} $I_{T_k}\left(\alpha_{T_{k-1}}\right)$ \remove{for any given strategy $\alpha_{T_{k-1}}$} by a suitable discretization of \eqref{eq:TVS_last} starting from $X_{T_{k-1}}$ and $I_{T_{k-1}}$.
\change{Thus, once collected the elements $\{X_{T_{k}}, I_{T_{k}}\left(\alpha_{T_{k-1}}\right)\}$,} the derivative price is given by:
\begin{equation}
\begin{aligned}
	    V_{T_{k-1}}\left(X_{T_{k-1}}, I_{T_{k-1}}\right)= \\ 
	     \mathbb{E}_{T_{k-1}}\left[D\left(T_{k-1},T_{k}\right) V_{T_{k}}\left(X_{T_{k}}, I_{T_{k}}\left(\alpha_{T_{k-1}}\right)\right) \mid X_{T_{k-1}}, I_{T_{k-1}}\right] \; ,
\end{aligned}
\label{eq:recursion2}\end{equation}
while the iteration starts from maturity date where the boundary condition is set equal to the payoff function:
\begin{equation}
    V_{T_m} = \Phi\left(I_{T_m}\right) \; .
\end{equation}

@@

\begin{proposition}\label{prop:bs}
When the underlying securities follow a Black and Scholes model with deterministic rates, the optimal strategy for a non-decreasing European payoff consists in minimizing the local drift function, independently of the current state $I_t$
\begin{equation}
    \alpha^*(t) \coloneqq \argmin_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t) \|} \; .
\label{eq:BS_optimal_strategy}
\end{equation}
Analogously, the optimal strategy for a non-increasing European payoff consists in maximizing the local drift function:
\begin{equation}
    \alpha^*(t) \coloneqq \argmax_\alpha \frac{\alpha \cdot \mu(t)}{\norm{\alpha \cdot \nu(t)}} \; .
\label{eq:BS_optimal_strategy_put}
\end{equation}
\end{proposition}
The absence of stochastic elements in Equation~\eqref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori} with no \change{numerical} simulation needed \change{to solve the control problem}; in fact\change{,} \remove{one can solve the optimization problem once for all $\forall t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities} \change{the functions $\mu(t)$ and $\nu(t)$ can be directly derived from the market quotes of the risky assets and of its derivative contracts, and then we can apply Equation~\eqref{eq:BS_optimal_strategy} $\forall t \in \mathcal{T}$  to retrieve the deterministic optimal strategy}. Once $\alpha^*$ is known, then one can price the payoff by the following BS formula
\begin{equation}
	V_0^{\text{BS}} = BS(F^{\text{TVS}}(0,T;\alpha^*), K, T, \bar{\sigma}, D(0,T;\zeta)) \; ,
\label{eq:bs_closed_solution}\end{equation}
where $F^{\text{TVS}}(t,T;\alpha)$ is the TVS forward curve defined by
\begin{equation}
	F^{\text{TVS}}(t,T;\alpha) = I_t \exp\left[\int_t^T\left(r(u)-\ell_{\alpha} \left(u\right)\right)du\right] \; ,
\label{eq:tvs_forward}\end{equation}
while $BS(F,K,T,\sigma,D)$ is the standard BS formula for a European option with forward curve $F$, strike $K$, time to maturity $T$, volatility $\sigma$ and discount factor $D$.

In Figure~\ref{fig:strategy_price_comparison} we provide a comparison among the option price obtained with the optimal strategy (BS$^*$) of Equation~\eqref{eq:BS_optimal_strategy} and those with \change{the below expert-based single-asset} \remove{other intuitive} strategies (S$_A$, S$_B$, S$_C$). Here we consider the case of an at-the-money call option with spot $I_0 = 1\, \text{EUR}$, maturity $T=5 \, \text{yr}$, target volatility $\bar{\sigma}=5\%$, and a nonnegative constraint on $\alpha$. The intuitive strategies are
\begin{itemize}
\item S$_A$: invest all in the asset with the maximum forward curve at maturity; \change{this strategy consists of selecting a static portfolio that is never rebalanced, such that the TVO moneyness is maximized;}
\item S$_B$: \remove{for each market pillar before maturity} \change{$\forall t \in \mathcal{T}$} invest all in the asset with minimum $\mu(t)$; \change{in this case we consider a fund manager who rebalances the weights $\alpha(t)$ over the life of the portfolio by selecting the asset with the highest expected return at a given point in time;} 
\item S$_C$: \remove{for each market pillar before maturity} \change{$\forall t \in \mathcal{T}$} invest all in the asset with minimum $\mu(t)/\|\nu(t)\|$\remove{.}\change{; the strategy is akin to S$_B$, with the exception being that the manager endeavors to maximize the expected return that has been adjusted to account for the target volatility mechanism.}
\end{itemize}
As the reader can observe, our strategy outperforms any other \remove{intuitive} \change{expert-based} one a practitioner might adopt \change{to face the control problem}.
\begin{figure}[t]
	\centering
	\includegraphics[width=4in]{strategy_price_comparison_grey.eps}
	\caption{Comparison of plain vanilla prices on the TVS by adopting different allocation strategies: the optimal Black and Scholes solution (BS$^*$) of Equation~\eqref{eq:BS_optimal_strategy} and \change{three expert-based single-asset} strategies (S$_A$, S$_B$, S$_C$)}
	\label{fig:strategy_price_comparison}
\end{figure}

@@

\subsubsection{Active \change{a}sset (or \change{b}ang\change{-b}ang) \change{s}olution}
A closed form solution to the minimization of the local drift correction \eqref{eq:BS_optimal_strategy} can also be derived in the common case that all costs of carry \change{$\mu$} are nonnegative\change{, so they represent true hedging costs and not benefits,} and the only constraint on portfolio weights is nonnegativity, which would mean a long-only strategy by the fund manager.

@@

Conversely, if we deal with a dynamic model for which the derivative contract $V$ depends on $X$ then the volatility versor term, namely the second one, in the $\max$ operator of Equation~\eqref{eq:HJB_general} is no longer zero and thus one must solve the entire control problem numerically. \change{Indeed, one can check by direct substitution that the natural generalization of \eqref{eq:HJB_solution}
\begin{equation}
 \alpha^*(X_t) = \argmin_\alpha \frac{\alpha \cdot \mu(X_t)}{\|\alpha \cdot \nu(X_t)\|}
\end{equation}
fails to satisfy \eqref{eq:HJB_general}.}

In \remove{this work}\change{what follows} we tackle the non-trivial case of a local volatility \remove{(LV)} model for the $S_t$-dynamics, such that $\nu_t = \nu(t,S_t)$. We have chosen the LV model since it is well known in financial literature and among practitioners \change{\cite{Gatheral2006}}.

\section{Reinforcement learning}\label{sec:RL}
As we have discussed in the previous sections, \change{the analytical solution described in Proposition \ref{prop:bs} crucially relies on the assumption that the underlyings evolve according to a Black and Scholes model, and on a monotonicity assumption on the payoff. Therefore,}
one must resort to numerical approaches to solve the stochastic control problem related to the TVS in the case of general payoffs or risky securities dynamics. The standard approach could be to use classical techniques based on \remove{backwards} \change{backward} recursion \eqref{eq:recursion}-\eqref{eq:recursion2} such as American Monte Carlo \cite{Longstaff2001}. However, their performances degrade exponentially as the dimension $n$ of the problem increases, making prohibitively costly finding the solution to the problem. In our contribution, we adopt a novel technique which is free from the curse of dimensionality and is gaining popularity in many scientific branches for solving stochastic optimal control problems: Reinforcement Learning (RL) \cite{Sutton2018}.

\remove{Reinforcement Learning} \change{RL} is a branch of Machine Learning which allows an artificial agent to interact with an environment through actions and observations in order to maximize some notion of cumulative reward \change{called value function}. In RL\change{,} the agent is not told which actions to take but instead, it \remove{must} \change{has to} discover by trial and error which are the behaviors  yielding the highest reward. This is obtained by updating the agent policy $\pi$ which is a mapping from the environment states to the set of actions. Thus RL is independent of pre-collected data as opposed to other Machine Learning techniques\change{, that have to be fed with a pre-existing dataset to learn from}. Because of its nature, RL has been successful in quantitative finance for solving control problems; among the most important RL applications in this field, we refer to \cite{Deng2017} as the pioneers in studying self-taught reinforcement trading problems, while to \cite{Halperin2020} and \cite{Kolm2019} for hedging derivatives with RL under market frictions.

@@

In this algorithm, we parameterize the agent policy with a feed forward neural network (FFNN), such that $\theta_j$ coincides with the hidden weights, $s_k$ with the \remove{inputs} \change{input neurons}, and $a_k$ with the output \change{ones}. In this way, we are dealing with a deterministic policy, \change{that maps each state directly to a single action without any additional randomness or probability distribution;} \remove{where its} \change{the map} functional form is given by the neural network. The parameters update is performed as follows: the agent collects a finite batch of experiences interacting with the environment in a set of episodes $\tau$, then the loss-\change{value} function is evaluated as
\begin{equation}
\change{L(\theta_j) =\hat{\mathbb{E}}\left[ \sum_{k=0}^{m-1} r_{k+1} | \theta_j \right]} \; ,
\label{eq:direct_policy_loss}\end{equation}
\begin{equation}
\hcancel[red]{\change{L(\theta_j) = D(0,T;\zeta)\hat{\mathbb{E}}\left[ \Phi(I_T(\alpha(\theta_j))) \right]}} \; ,
\end{equation}  
where the expectation $\hat{\mathbb{E}}[\dots]$ indicates the empirical average on the batch. Then the parameters are updated by plugging the policy into a stochastic gradient ascent algorithm \change{\cite{Sutton2018} with the aim of maximizing Equation~\eqref{eq:direct_policy_loss}}. \remove{The choice of Equation~\eqref{eq:direct_policy_loss} is justified by the fact that, when the algorithm will find the optimal set $\theta^*$, then $L(\theta^*)$ will be a good proxy of the optimal option price.}

@@

\change{As the PPO algorithm is more sophisticated than the previous one, we have decided to test its performance by means of two different reward functions. The former is the same as defined by Equation~\eqref{eq:reward1}, while the latter is given by}
\remove{The second reward function we have defined is}
\begin{equation}
 	r_{k+1} \coloneqq  \gamma^{k}[V_{\text{BS}}(T_{k+1})-V_\text{\text{BS}}(T_k)] \; ,
 \label{eq:reward2}\end{equation}
where $\gamma \in[0,1]$ is an hyper-parameter of the PPO, while $V_{\text{BS}}(T_k)$ is a proxy of the residual option price defined by
\begin{equation}
\begin{aligned}
	V_{\text{BS}}(T_k) &\coloneqq BS(F^{\text{TVS}}(T_k,T;\alpha^*_{\text{BS}},K),K,T-T_k,\bar{\sigma},D(T_k,T;\zeta)) \; ,\\
	V_{\text{BS}}(T_0)&=0 \; , 
\end{aligned}
\end{equation} 
with $\alpha^*_{BS}$ the BS optimal strategy \eqref{eq:BS_optimal_strategy} evaluated at the state $s_k$. In this form the agent actions are hidden inside the term $I_t$ used to compute the TVS forward curve $F^{\text{TVS}}$ defined by Equation~\eqref{eq:tvs_forward}. In this case, the reward function does not suffer of nil values for $0<T_k<T$: the RL agent always gets a feedback from the environment for its choices.  \remove{The} \change{Notice that the} hyper-parameter $\gamma$ plays the role of a discount factor in the sense that, as $\gamma$ approaches to zero, the RL agent will tend to maximize immediate rewards while neglecting possible larger rewards in the future.
If we take the cumulative reward per episode and set the PPO parameters\footnote{We refer to \cite{Schulman2016} for a more detailed description for the generalized advantage estimation parameter $\lambda$.} $\gamma =\lambda =1$ we obtain
\begin{equation}
	R(\tau) = \sum_{k=0}^{m-1} r_{k+1} \underset{\gamma=\lambda =1}{=} \sum_{k=0}^{m-1}[V_{\text{BS}}(T_{k+1})-V_\text{\text{BS}}(T_k)] = V_{BS}(T) = (I_T-K)^+ \; ,
\end{equation} 
which is equal to the intrinsic value of the option. This result does not depend on the definition of $V_{\text{BS}}$  $\forall T_k<T$, but we conjecture that the closer $V_{\text{BS}}$ is to the value function, the easier the agent is in learning.
